---
title: "gov51_ps2"
format: html
editor: visual
---

## Gov 51 problem Set 2

By Sage Piekarski

Part 1 Question 1

<https://github.com/sagepiekarski-ai/gov51-ps2.git>

```{r}

#loading tidyverse
##install.packages("tidyverse")
library(tidyverse)

#read csv
gay <- read.csv("Data/Raw/gay.csv")

#calculate unique values
unique_value <- length(unique(gay$study))
print(unique_value)

#identify treatment conditions
treatment_cond <- unique(gay$treatment)
print(treatment_cond)

```

Q1.1 There are 69,592 observations in gay.csv. There are 2 unique values in study. The five possible treatment conditions are "No contact", "Recycling Script by Gay Canvasser", "Same-sex Marriage Script by Gay Canvasser", "Recycling Script by Straight Canvasser", and "Same-sex Marriage Scirpt by Straight Canvasser".

```{r}

#filter gay
gay_clean <- gay %>% 
  filter(study == 1, 
        wave == 1)
ssm <- gay_clean %>% 
  group_by(treatment) %>% 
  summarize(ssm_mean = mean(ssm, na.rm = TRUE))
print(ssm)

```

Q1.2 The group that is treated with a recycling script by a gay canvasser shows the highest average support.

```{r}

#calculate sd
sd <- gay_clean %>% 
  group_by(treatment) %>% 
  summarize(ssm_sd = sd(ssm, na.rm = TRUE))
print(sd)
```

Q1.3 Yes, for the most part, the standard deviations are roughly similar across groups with the sd values ranging from 1.67 to 1.70. In a randomized experiment, standard deviations are very important because they indicate the spread of the data and to what extent values are from the mean. If the standard deviation is all relatively similar and small, then this indicated our randomized experiment is working properly and observed differences can be attributed to the treatment.

```{r}

#create a bar plot of the mean ssm by treatment
ggplot(ssm, aes(x = treatment, y = ssm_mean)) +
  geom_col(fill = "cornflowerblue") +
  labs(title = "Mean Same Sex Marriage by Treatment Group",
       x = "Treatment Group",
       y = "Mean SSM Score") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

Question 1 Part 2

```{r}

#filter gay_clean more
gay_clean2 <- gay_clean %>% 
  filter(treatment == "No Contact"| treatment == "Same-Sex Marriage Script by Gay Canvasser")

#calculate Mean & SD
summarized <- gay_clean2 %>% 
  group_by(treatment) %>% 
  summarize(ssm_mean2 =  mean(ssm, na.rm = TRUE),
            ssm_sd2 = sd(ssm, na.rm = TRUE))
print(summarized)

#calculate Standardized difference in means
#numerator
(3.025195 - 3.042764)
#find denominator
1.666099^2 + 1.680540^2
5.600101 / 2
sqrt(2.80005)
0.017569 / 1.673335

```

Q1.5. The standardized difference in means is -0.01049939. The mean and SD for each group is located above in the table. This difference is significantly below 0.25 which leads me to believe that in fact their is good balance and the two groups are very similar at baseline.

```{r}

#filter gay
gay_clean3 <- gay %>% 
  filter(study == 1, 
        wave == 2,treatment == "No Contact"| treatment == "Same-Sex Marriage Script by Gay Canvasser")

#calculate Mean & SD
summarized2 <- gay_clean3 %>% 
  group_by(treatment) %>% 
  summarize(ssm_mean3 =  mean(ssm, na.rm = TRUE),
            ssm_sd3 = sd(ssm, na.rm = TRUE))
print(summarized2)

#difference in means
3.139489 - 3.039615

```

Q1.6 the difference in means here is 0.099874. Since the groups were balanced at baseline, a difference at wave 2 suggests that this difference was because of the treatment rather than any outside factors in the groups. In broader terms, the implementation of a gay canvasser had a small positive impact on support for same sex marriage in reference to the no contact group.

**Question 2**

```{r}

#load more data
gayreshaped <- read.csv("Data/Raw/gayreshaped.csv")
ccap2012 <- read.csv("Data/Raw/ccap2012.csv")
```

Q2.1

gayreshaped has 11,948 observations. Its columns are study, treatment, therm1, therm 2, therm3, and therm4.

ccap2012 has 43,998 observations. Its column names are x, caseid, and gaytherm.

```{r}
#summary statistics for gayreshaped therm1
mean(gayreshaped$therm1[gayreshaped$study == 1], na.rm = TRUE)
median(gayreshaped$therm1[gayreshaped$study == 1], na.rm = TRUE)
sd(gayreshaped$therm1[gayreshaped$study == 1], na.rm = TRUE)

#summary statistics for ccap2012 gaytherm
mean(ccap2012$gaytherm, na.rm = TRUE)
median(ccap2012$gaytherm, na.rm = TRUE)
sd(ccap2012$gaytherm, na.rm = TRUE)

side_by_side <- data.frame(Dataset = c("Study 1 (therm1)", "CCAP 2012 (gaytherm)"),
                           Mean = c(mean(gayreshaped$therm1, na.rm = TRUE),
                                    mean(ccap2012$gaytherm, na.rm = TRUE)), 
                           Median = c(median(gayreshaped$therm1, na.rm = TRUE),
                                    median(ccap2012$gaytherm, na.rm = TRUE)), 
                           SD = c(sd(gayreshaped$therm1, na.rm = TRUE),
                                    sd(ccap2012$gaytherm, na.rm = TRUE)))
side_by_side
                                    
                                    

```

Q2.2 From this side by side table, I can see that the mean, median, and sd of the data from gayreshpaed and CCAP2012 look very similar. In comparing the two, CCAP2012 has slightly larger values of mean, median, and SD compared to gayreshaped, but they are all very similar.

```{r}

#create side by side histograms
gayreshaped_study1_therm1 <- gayreshaped %>% 
  filter(study == 1) %>% 
  select(therm1) %>% 
    rename(therm = therm1) %>% 
  mutate(source = "Lacour")


ccap2012_gaytherm <- ccap2012 %>% 
  filter(!is.na(gaytherm)) %>% 
  select(gaytherm) %>% 
  rename(therm = gaytherm) %>% 
    mutate(source = "CCAP")


side_by_side_data <- bind_rows(gayreshaped_study1_therm1, ccap2012_gaytherm)

ggplot(side_by_side_data, aes(x = therm)) + 
  geom_histogram(breaks = seq(0, 100, by = 5), 
                 color = "cornflowerblue", 
                 fill = "white") +
    facet_wrap(~source) +
  labs(title = "LaCour Study 1 baseline (therm1) vs the CCAP data (gaytherm)",
       x = "Feeling Thermometer Scores (0-100)",
       y = "Count")
  
```

Q2.3 I DONT KNOWW

**Question 2 Part 2**

```{r}

#filter new data
study1_control <- gayreshaped %>% 
  filter(study == 1, treatment == "No Contact") %>% 
  select(therm1, therm2)

study2_control <- gayreshaped %>% 
  filter(study == 2, treatment == "No Contact") %>% 
  select(therm1, therm2)

#calculate study 1 values
study1_correlation <- cor(study1_control$therm1, study1_control$therm2, use = "complete.obs")
study1_correlation
study1_sd <- sd(study1_control$therm2 - study1_control$therm1, na.rm = TRUE)
study1_sd
#calculate study 2 values
study2_correlation <- cor(study2_control$therm1, study2_control$therm2, use = "complete.obs")
study2_correlation
study2_sd <- sd(study2_control$therm2 - study2_control$therm1, na.rm = TRUE)
study2_sd

#put side by side
side_by_side_results <- data.frame(Study = c("Study 1", "Study 2"),
                                   Correlation = c(study1_correlation, study2_correlation),
                                   StandardDeviation = c(study1_sd, study2_sd))
side_by_side_results
```

Q2.4 If we first look at Study 1, we will see that there is a correlation of 0.998 which is quite a bit higher than the typical for real feeling thermometer data of 0.95 to 0.97. In other words, from therm1 to therm2 participants' responses were likely very similar. Study 2's correlation, 0.973 tends to look more similar to what we would expect with the value not far outside of the 0.95 to 0.97 range. Examining the standard deviations of the within person change, we can see that study 1's standard deviation of 1.988 is quite small while study 2's 6.606 seems to be more realistic and account for within person differences. This information all leads me to believe that study 2 is likely showing a more realistic and plausible pattern.

```{r}

#create two side by side scatterplots
combined_studies <- bind_rows(study1_control %>% mutate(Study = "Study1"),
                              study2_control %>% mutate(Study = "Study2"))

ggplot(combined_studies, aes(x = therm1, y = therm2)) +
  geom_point(color = "cornflowerblue") +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~Study) +
  labs(title = "Study 1 vs Study 2 Control Group",
       x = "Therm 1",
       y = "Therm 2")
```

Q2.5 Study 2 definitely looks more like real longitudinal data. If you take a look at the right scatter plot all of the data is incredibly close to the 45 degree line and doesn't mimic the actual variation you would probably see. As a result, study 2 looks more realistic because it shows that more widespread measure of spread that study 1 lacks. Study 1 also seems to have that "t" that we talked about in lecture in the middle right by 50.

**QUESTION 2.6**

**Part 3**

```{r}

#filter gay reshaped
gayreshaped_clean2 <- gayreshaped %>% 
  filter(study == 2)

mean(gayreshaped_clean2$therm1[gayreshaped_clean2$treatment == "No Contact"], na.rm = TRUE)
sd(gayreshaped_clean2$therm1[gayreshaped_clean2$treatment == "No Contact"], na.rm = TRUE)


mean(gayreshaped_clean2$therm1[gayreshaped_clean2$treatment == "Same-Sex Marriage Script by Gay Canvasser"], na.rm = TRUE)
sd(gayreshaped_clean2$therm1[gayreshaped_clean2$treatment == "Same-Sex Marriage Script by Gay Canvasser"], na.rm = TRUE)

#create side by side histograms
histo_study1 <- gayreshaped %>% 
  filter(study == 1) %>% 
  select(therm1) %>% 
  rename(therm = therm1) %>% 
  mutate(Study = "Study1")

histo_study2 <- gayreshaped %>% 
  filter(study == 2) %>% 
  select(therm1) %>% 
  rename(therm = therm1) %>% 
  mutate(Study = "Study2")

combined_histo <- bind_rows(histo_study1, histo_study2)

ggplot(combined_histo, aes(x = therm)) +
  geom_histogram(breaks = seq(0, 100, by = 5),
                 color = "cornflowerblue",
                 fill = "white") +
  facet_wrap(~Study) +
  labs(title =  "Study 1 baseline (therm1) and Study 2 baseline (therm1)",
       x = "Feeling Thermometer Scores (0-100)",
       y = "Count")
                 
```

Q3.2 Visually it appears as though there are a number more counts observed in study 1 than in study 2. In both histograms we observe spikes at 0, 100, and around 50. These histograms look very similar to the two histograms we generated earlier in the problem set.

```{r}

#filter gay

q3.3 <- gayreshaped %>% 
  filter(study == 2, treatment == "No Contact" | treatment == "Same-Sex Marriage Script by Gay Canvasser") 

#calculate Mean & SD
summarized3 <- q3.3 %>% 
  group_by(treatment) %>% 
  summarize(q3.3_mean =  mean(therm1, na.rm = TRUE),
            q3.3_sd = sd(therm1, na.rm = TRUE))
print(summarized3)

#calculate Standardized difference in means
#numerator
(59.35057 - 57.88612)
#find denominator
28.41578^2 + 28.69655^2
1630.949 / 2
sqrt(815.4745)
1.46445 / 28.55651
  
```

Q3.3 The mean and standard deviations are listed in the table above. The standardized difference in means is 0.05128253. The groups do seem to be balanced with their means and SD's looking close to eachother and the standardized difference in means being low.

```{r}

#Q3.4
summarized4 <- q3.3 %>% 
  group_by(treatment) %>% 
  summarize(q3.4_mean =  mean(therm2, na.rm = TRUE))
print(summarized4)

61.20311 - 57.63523

```

Q3.4 The difference in post period means is 3.56788.

```{r}

#calculate the standard error
#gay canvasser
61.20311^2
sum(q3.3$treatment == "Same-Sex Marriage Script by Gay Canvasser") 
#no contact
57.63523^2
sum(q3.3$treatment == "No Contact")

#gay canvasser
3745.821/1238

#no contact
3321.82/1203

3.025704 + 2.76128

sqrt(5.786984)

#calculate confidence interval
61.20311 - 57.63523
3.56788 + 1.96*2.405615
3.56788 - 1.96*2.405615
```

Q3.5 the standard error is 2.405615. The confidence interval is \[-1.147125, 8.282885\]. Yes, this confidence interval does include zero. This tells us that our output is likely not statistically significant and we may observe no correlation in our output.

```{r}

#calculate correlation
study2_correlation2 <- cor(q3.3$therm1, q3.3$therm2, use = "complete.obs")
study2_correlation2
study2_correlation3 <- cor(q3.3$therm1, q3.3$therm3, use = "complete.obs")
study2_correlation3
study2_correlation4 <- cor(q3.3$therm1, q3.3$therm4, use = "complete.obs")
study2_correlation4

```

Q3.6 Overtime, the calculations decay from therm1, therm2 and therm1, therm3. However, from therm1, therm3, to therm1, therm 4 the correlation actually rises a bit. All in all, observing the correlation from therm1, therm2, to therm1, therm4, the correlation does fall slightly. Ultimately, these correlations all stay fairly close together, but this tells us that in reality stability is not perfect and

**Q3.7 MUST DOOOO**

**Part 4**

```{r}

#cleandata
q4.1 <- gayreshaped %>% 
  filter(study == 1, treatment == "No Contact")
  
#Fit a linear regression

fit_s1 <- lm(therm2 ~ therm1, data = q4.1)
summary(fit_s1)
```

Q4.1 the slope, or beta1, is 0.992454. The R-squared is 0.9952 and the residual standard error is 1.976.

```{r}

#cleandata
q4.2 <- gayreshaped %>% 
  filter(study == 2, treatment == "No Contact")
  
#Fit a linear regression

fit_s2 <- lm(therm2 ~ therm1, data = q4.2)
summary(fit_s2)

```

Q4.2 the slope, or beta1, is 0.968662. The R-squared is 0.9476 and the residual standard error is 6.547. If we are examining these two in the context of fabricated data, study 1 definitely seems to be more fabricated than study 2. With a slope and r-squared value (0.992454 and 0.9952) of approximately 1, it seems likely that this data could be false. It seems as though study 2 is more of what we would expect realistic data to look like with a R-squared value and a slope value both more significantly less that one but not that far off.

```{r}

#cleandata
q4.3 <- gayreshaped %>% 
  filter(study == 2, 
  treatment %in% c("No Contact", "Same-Sex Marriage Script by Gay Canvasser")) %>% 
  mutate(treated = ifelse(treatment == "Same-Sex Marriage Script by Gay Canvasser", 1, 0))
  
#Fit a linear regression
fit_s3 <- lm(therm2 ~ treated, data = q4.3)
summary(fit_s3)



```

Q4.3 Now the slope, or beta1, is 3.5679. The standard error is 1.2423. The t-statistics is 2.872. The p-value is 0.00412. Comparing beta1 to the difference in means from question 3.4, we are reminded that the difference in means was 3.56788, or approximately the same exact value of the observed beta1 value. beta not estimates to 57.6352. We know the p-value is 0.00412, so looking at a an alpha = 0.05 level, we can conclude that the outcome is statistically significant because the p-value is less than the alpha.

```{r}

#Fit a linear regression
fit_s4 <- lm(therm2 ~ treated + therm1, data = q4.3)
summary(fit_s4)

```

Q4.4 The coefficient on treated is 2.458857. Its standard error is 0.341052. The t-statistic is 7.21 and its p-value is 7.76e-13. In comparison to the results from question 4.3, the beta1 coefficient did decrease by a fair amount from 3.5679 to 2.458857. The standard error did decrease quite a bit from 1.2423 to 0.341052 indicating the coeffficient estimate became more precise. Controlling for baseline attitudes definitely improves precision because by including these controls, we can account for any pre existing differences in atttidues. For instance, some people may begin super accepting of gay individuals while others may be very much against them. By accounting for where people start in terms of gay acceptance, we can more clearly see the impact of the treatment group which makes the precision of the study much more enhanced.

```{r}

#use predict
q4.5 <- data.frame(treated = c(1, 0),
                   therm1 = c(50, 50))

q4.5_pv <- predict(fit_s4, newdata = q4.5)
q4.5_pv

52.47638 - 50.01752

```

Q4.5 The R-squared value from Q4.3 is 0.003857 and the R-squared value from question 4.4 is 0.925. Adding therm1 as in Q4.4 increases the R-squared so dramatically because the R-squared is a measure of the percent of variance explained by the model. In the first regression, the R-sqaured is lower because it suggestes there is a lot of variance not explained by the mode. In the second regression, we control for baseline attitudes, thus, the percent of variance explained by the model is significantly higher. This does indicate that a fair amount of the variation in follow-up attitudes is explained by where people started versus the treatment. The predicted difference in their wave 2 attitudes is 2.45886.
